A Compound Decision Approach to Covariance Matrix Estimation

### 2021-12-01 response_reviews.tex

\title{Response to reviews of BIOM2021236M}
\author{}
\date{}

\begin{document}
\maketitle

\section{Response to the Associate Editor}

Thank you for your very valuable comments. We respond to each point below.

\begin{enumerate}
\item \emph{As a journal, Biometrics presents papers that focus on the development of new methods and results of use in biosciences. As of now, the application seems more an afterthought, the use of a convenient (easily available) dataset to fit the scope of the journal. I believe that the authors should motivate their approach from the gene expression data or other bio-related applications starting from the Introduction.}

  We have revised our introduction and have also completely redone our data analysis to study a question about gene network inference that arose during some of our previous collaborative work. We were involved in an analysis of bulk RNA-seq expression from mouse brain in a study of the neurogenomic response to social intrusion (Saul et al., 2017). We have now used some of these data to study gene networks in different regions of the mouse brain; \textbf{see Section 5}.

\item \emph{It is not really intuitively clear why the vectorization works, despite the approach not taking into consideration the structure of the covariance matrix. Also, the manuscript proposes a series of ad-hoc expedients to implement the method. See the clustering-based exemplar algorithm and replacing non-positive eigenvalues with a chosen positive value $c$ in the positive-definite correction. These expedients do not seem to be completely justified by the proposed separable compound-decision theory framework, and their effects on the estimation are unclear. A referee also commented on the impossibility to follow closely the implementation of the EM algorithm. In short, while I appreciate that the theoretical justification for the asymptotic efficiency of the proposed vectorization is not clear, the current implementation reads more like a clever algorithm than a well-founded statistical method. The authors should justify all the steps of the algorithm and provide an assessment of their contribution to its overall performance.}

  The success of our vectorization perspective is indeed surprising. In our revision we have made several attempts to contextualize this finding.
  \begin{enumerate}

  \item In \textbf{Section 2.1} we now note that measuring the success of covariance matrix estimation using Frobenius risk does not incentivize positive-definiteness. In other words, there can exist estimators of $\bs{\Sigma}$ that have low Frobenius risk but which are not positive-definite. So the fact that vectorization works is in part a consequence of the risk function.

  \item In \textbf{Section 2.2} we note that several existing shrinkage methods can already be interpreted as vectorized methods. The sample covariance matrix, the linear shrinkage method of Ledoit and Wolf (2004), and the adaptive thresholding method of Cai and Liu (2011). The adapative thresholding method is also not guaranteed to be positive-definite.

  \item In \textbf{Section 4.4} we now study one possible reason why vectorization may work better than existing rotationally invariant procedures that shrink only eigenvalues but use sample eigenvectors in their estimators. Since sample eigenvectors can be poor estimators of population eigenvectors in high-dimensions, we hypothesized that our vectorized approach would be better in simulation settings where the sample and population eigenvectors were more different. This appears to be the case; \textbf{see Figure 3}.
  \end{enumerate}

  We agree that projecting our final estimator on to the space of positive-definite matrices is not warranted by the compound decision formalism. Indeed, our uncorrected estimator still works very well in our simulation settings, and furthermore the (uncorrected) oracle vectorized estimator can be dramatically better than the oracle rotationally invariant estimator; \textbf{see Figure 2}. Nevertheless, it does appear that in finite samples, positive-definiteness correction can improve performance. We mention in the \textbf{Discussion} that a non-separable class of estimators that can incorporate this constraint is worth exploring in the future.

  In our new \textbf{Section 3.3} we have also clarified our EM updates and have added details about iimplementation.

  In short, in our paper we explore a new perspective on covariance matrix estimation, a vectorization perspective founded on a compound decision theoretic interpretation of the Frobenius risk. We define a new class of estimators, derive the optimal vectorized estimator within this class, and estimate the optimal estimator using approaches from the compound decision literature. Finally, we argue that these estimators work well for covariance matrices whose sample and population eigenvectors are very different.

\item \emph{I agree with one of the referees that the title should be re-thought, since compound decision theory or compound decisions appear quite relevant to the approach, more than the empirical Bayes aspect of the paper.}

  We have changed our title following your suggestions.

\item \emph{on page 4, it appears to me that the choice of the risk function $\hat{R}$ does not correspond to the Frobenius risk. The rationale for choosing this risk function should be provided. It somehow appears the choice is dictated by Ledoit et al (2004) and Proposition 1, and not by independent first principles. If so, it should be explicitly stated before.}

  We have clarified in \textbf{Section 2.2} that the choice $\hat{R}$ follows directly from first principles.
  \begin{enumerate}
  \item In (4) we restrict ourselves to the class of linear vectorized estimators of $\bs{\Sigma}$.

  \item In Proposition 1 we show that $\hat{R}$ is an asymptotically unbiased estimate of the Frobenius risk for any fixed $\bs{\Sigma}$ and any fixed linear estimator.

  \item A natural procedure is to minimize the empirical risk $\hat{R}$ over the class of linear estimators.

  \item In Proposition 2 we show that the estimator obtained after empirical risk minimization is very closely related to the Ledoit and Wolf (2004) estimator.
  \end{enumerate}

\item \emph{Figures 1 and 2 do not allow to appreciate uncertainty. Perhaps a table would be better.}

  We now give exact numerical results for all simulations in our \textbf{Supporting Information}.

\item \emph{The manuscript requires careful proofreading and improvements. in the structure of the presentation and the visualization of the results. Non-exhaustive examples are reported below:}

@@

\emph{- p. 2 “Other common structured methods assume that the covariance matrix is banding” $\to$ banded?}

This has been edited to ``banded''.

\emph{- p. 2 “yet can still outperform the sample covariance matrix” What does this mean?}

It means that these unstructured model have been shown to have lower estimation risk comparing to sample covariance matrix. To make it clear, we have changed to ``yet still have less estimation error comparing to the sample covariance matrix''.

@@

\emph{- p. 16 “Finally, we have so far assumed that our data multivariate normal” $\to$ are normal}

This has been fixed.

\emph{- In Equations (1) and (2) the arguments of the risk functions are $\theta$ and $\delta$. However, on page 5 the arguments of $\hat{R}$ are $\beta_{S}$, and $\beta_{I}$.}

We now use $\hat{R}_L$ to denote the risk estimate for the class of linear decision rules. We also make the distinction between $\hat{R}_L$ and $R$ more explicit in \textbf{Proposition 1}.
\end{enumerate}

\section{Response to Referee 1}

Thank you for your positive comments, and your careful reading and extremely helpful suggestions. We respond to them below.

\subsection{Response to referee section ``CorShrink''}

Thank you so much for bringing our attention to Dey and Stephens (2018). It is indeed extremely closely related to our work. We now discuss connections to their work in several places.

\begin{enumerate}
\item In the last paragraphs of \textbf{Section 3.2} we note that the main difference between our approaches is that Dey and Stephens (2018) do not shrink the variances while we do, as our $\hat{t}_d$ estimates $\sigma_j$ using its supposed posterior expectation assuming a prior distribution of $\hat{G}_d$. Furthermore, our Proposition \ref{prop:bayes risk} shows that our shrinkage approach can be interpreted as viewing the triples $(\sigma_j, \sigma_k, r_{jk})$ as if they were drawn from the trivariate prior $G_{od}$, which allows for possible dependencies between $r_{jk}$ and the $\sigma_j$. For example, model 2 in our simulations in Section \ref{numerical results} considers a covariance matrix where entries with larger standard deviations also tend to have larger correlations. In contrast to the method of Dey and Stephens (2018), our approach can learn this structure if it exists and take advantage of it for more accurate estimation. We demonstrate this in Model 2 in our simulations.

\item We have added CorShrink to our simulation studies for estimating both covariance matrices (\textbf{Figure \ref{fig:sim1_frobenius}}) and correlation matrices (\textbf{Table 3 in the Supporting Information}). We found that for estimating covariance matrices, our approach could outperform CorShrink in many cases, likely because it was able to accurately shrink the standard deviations. In the \textbf{Supporting Information} we also compare MSG and MSGCor to CorShrink for estimating correlation instead of covariance matrices. We find that in that case, CorShrink performs slightly better than MSG and MSGCor except in our Model 2, likely because the additional flexibility that our method trades off lower bias for higher variance.

\end{enumerate}

\subsection{Response to referee section ``Separable rules, symmetry and the diagonal''}

\begin{enumerate}
\item \emph{While reading the paper I was a bit confused by the fact that (R1) treats diagonal entries and off-diagonal in the same way. It would feel more natural to me to define a separable estimator as in (R1) for $j\neq k$ \ldots In the Multivariate Gaussian case this is not an issue, since $X_{\cdot j}\neq X_{\cdot k}$ for $j\neq k$ with probability 1. But the proposed approach would also apply with minor modifications to e.g., discrete measurements (with known likelihood) in which case $X_{\cdot j}\neq X_{\cdot k}$ could happen with non-zero probability. Furthermore, while this issue is ignored for most of the paper, diagonal and off-diagonal entries are treated in a different way at the end of Section 2.3 (in the context of the clustering based algorithm). I believe it would be valuable to make the above distinction more explicit!}

  Thank you for this excellent suggestion. We completely agree and have now redefined our class of separable estimators to be
  \[
  \mathcal{S} = \{\bs{\delta} : \delta_{kj} = \delta_{jk} = t(\bs{X}_{\cdot j}, \bs{X}_{\cdot k}),\, 1\leq k<j\leq p,\quad 
   \delta_{jj} = \widetilde{t}(\bs{X}_{\cdot j}),\,j=1,\ldots,p)\},
 \]
 so that the on- and off-diagonal entries are treated separately. In \textbf{Section \ref{sec:class}} we derive the optimal estimator in this class. This estimator turns out to be the same as the estimator we actually implemented in our previous draft, but redefining $\mathcal{S}$ makes this much more rigorous.

\item \emph{A related question I had concerns the symmetry of the estimator. Does the estimator satisfy the symmetry property $\hat{\sigma}_{jk} =\hat{\sigma}_{kj}$? Similarly, the reader may wonder whether the estimation procedure enforces symmetry in the prior $G$ in (6) with respect to $\sigma_j,\sigma_k$.}

  In our new implementation we now explicitly enforce symmetry. We construct the support points of our estimated discrete prior to be symmetric and we also force the estimated masses at these points to be symmetric as well. See \textbf{Section \ref{implementation}} for details.

\item \emph{The paper emphasizes the Multivariate Gaussian problem. In that case, the sample covariance is sufficient, and so, it would seem more natural to me to define a separable rule, as one that satisfies,
$$\hat{\sigma}_{jk} = \bar{t}(X_{\cdot j}^TX_{\cdot k}),\quad \bar{t}: \bb{R}\to \bb{R}_{+}$$
instead of (R1). I am wondering why the authors chose to use (R1) instead.}

We chose our definition of a separable estimator because it is more general than $\bar{t}(\bs{X}_{\cdot j}^\top \bs{X}_{\cdot k})$. We actually initially tried this class of estimators but found that its performance was worse than the estimators we proposed in the manuscript. Intuitively, the reason is that the statistic $X_{\cdot j}^\top X_{\cdot j}$ is not sufficient for the parameter $\sigma_{jk}$.

\item \emph{If I understand correctly, one of the benefits of estimator (3) is that it works without specific assumptions on the noise distribution. Instead the approach here appears to more heavily rely on a known likelihood (and specifically a multivariate Gaussian likelihood). It would be great if some simulations could be conducted under misspecification, wherein the data is not multivariate Gaussian, but the approach is applied assuming multivariate Gaussian data.}

  We now study the performance of our proposed method when the multivariate Gaussian assumption is violated. \textbf{Tables 4 and 5 in the Supporting Information} show that our procedures can still provide relatively accurate estimates for uniform or negative binomial data.

\end{enumerate}

\subsection{Response to referee section ``Compound decision theory or empirical Bayes?''}
\begin{enumerate}
\item \emph{The result following equation (6) could be presented separately as a Theorem (or Proposition). It should be highlighted that part of the statement is that the estimator in (R1) that optimizes (R2) depends on all the $\sigma_{j}$ and $\rho_{jk}$.}

  We have followed your suggestion and added \textbf{Proposition \ref{prop:bayes risk}} along with a discussion after it highlighting the fact that the optimal estimator depends on all the $\sigma_{j}$ and $\rho_{jk}$.

\item \emph{The term ``compound decision theory'' or ``compound decisions'' etc. deserves to be in the title!}

  Thank you for the suggestion! We have revised the title accordingly.

\item \emph{On the other hand, I would downemphasize the empirical Bayes aspect of the paper (and my preference would be to even remove it from the title). To me, an empirical Bayes analysis (especially since here the goal is to perform well, in a frequentist sense, in terms of (R2)) has the following flavor: ``Let us mimic what a Bayesian would do, when we do not know the true prior, and in a way that we can get frequentist guarantees as well''. However, I do not think a Bayesian would start approaching this problem by first positing (6) and then using (8)2. Instead a more natural Bayesian approach would be to set a prior for the full covariance matrix $\Sigma$, e.g., as in Berger et al. [2020] and references therein. I would restrict the discussion of empirical Bayes to just saying that model (6) appears formally when optimizing over separable rules (R1). To approximate the unknown optimal rule, one follows an empirical Bayesian ``Generalized Maximum Likelihood'' approach with respect to the pairwise composite likelihood.}

  This is an excellent point and we have done exactly as you suggested. We now only refer to empirical Bayes ideas as a method of estimating the optimal compound decision rule.

\end{enumerate}

\subsection{Response to section ``Other remarks''}
\begin{enumerate}
\item \emph{It would be of great service to the community if the authors could provide code implementing the method and reproducing the results!}

  We now note at the \textbf{end of the Introduction} that our code is available on Github and give the link.

\item \emph{Related to the above point, based on the description in the manuscript it would not be possible for me to fully reproduce the results. For example, does the actual implementation use the EM algorithm or an interior point convex solver? If the EM algorithm is used, how many steps are taken or what is the termination criterion? And why is the EM algorithm used (since for these empirical Bayes problem it typically converges extremely slowly compared to interior point solvers, as carefully demonstrated by Koenker and Mizera [2014])? On the other hand, if an interior point solver is used, which one?}

  We have revised our description of our implementation in \textbf{Section \ref{implementation}} to make our EM algorithm more clear. We aso note that instead of EM, we attempted to maximize the composite likelihood \eqref{Gnd hat} using the interior point solver MOSEK, as advocated by Koenker and Mizera (2014), but ran into computational difficulties because for even moderate sample sizes $n$, the values of the likelihoods we needed to calculate could be extremely small at certain support points. These small values caused problems for the R wrapper for MOSEK. Our implementation of the EM algorithm was more robust to these small density values.

\item \emph{First paragraph of page 3, ``they modify only the sample eigenvalues and not the sample eigenvectors'': The situation here is a bit more nuanced, since e.g., sometimes using the sample eigenvectors is provably the optimal thing to do (e.g., if one constrains themselves to rotationally invariant estimators as in Bun et al. [2016] and references therein)}

  Thank you for this reference. We now explicitly distinguish our proposed method from rotationally invariant ones and note in our \textbf{Introduction} that while rotationally invariate estimators perform extremely well, Bun et al. (2016) showed that the eigenvectors of any rotationally invariant estimate must be the same as those of the sample covariance matrix. This can be problematic because sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate  (Mestre, 2008). This suggests that unstructured but non-rotationally invariant covariance matrix estimators may be worth exploring, as these can modify both the sample eigenvectors as well as the eigenvalues. Our contribution is to propose and study one such class of non-rotationally invariant estimators.

  Incidentally, our new simulations in \textbf{Figure \ref{fig:eigenvectors}} suggest that the simulation settings where our estimators can outperform existing methods are exactly those where the sample and population eigenvectors differ most.

\item \emph{Page 5, ``it is straightforward to show that'': The ``straightforward'' could be replaced by a short proof in the appendix.}

  We have rewritten this as \textbf{Proposition \ref{prop:Rhat}} and have added a proof to the Supporting Information.

\item \emph{Also could it be that the risk estimate is not exactly unbiased (because of dividing by $n$ and not $n-1$ in the definition of $\hat{\Delta}_{jk}^2$)?}

  You are absolutely right, thank you for catching this. We have changed our text to mention that the risk estimate is only asymptotically unbaised, and we show this in the Supporting Information.

\item \emph{Where is the sample covariance $\bs{S}$ defined?}

  We now define $\bs{S}$ in the first paragraph of \textbf{Section \ref{sec:connections}}.

\item \emph{Section 2.5, ``Projections in terms of other matrix norms are also possible''. Is this actually being done in the implementation?}

  We did not implement this and so have removed this comment from the manuscript.

\item \emph{Else perhaps remove this sentence.}

This has been removed.

\end{enumerate}

\section{Response to Referee 2}

Thank you for your positive comments. We respond to them below.

\begin{enumerate}
\item \emph{In the numerical experiments and the real data analysis, what was the size of the grid that was used? I may have missed it but I did not see it mentioned in the caption for figure 2.}

  We have now added a separate \textbf{Section \ref{optimalK}} that specifies the choice of grid size for our proposed method.

\item \emph{In the numerical experiments, the dimension p goes as far as 200. I believe it will be useful to consider a setting where $n = 100$ and $p = 1000$.}

  We have added a simulation with $p = 1000$ in the \textbf{Supporting Information} and in \textbf{Section \ref{sec:accuracies}} note that while our methods perform well in our Models 1 through 4, other methods are better in Models 5 and 6 for large $p$. This is likely due to the fact that our class of estimators may not be be optimal in these settings because the population eigenvectors in these models appear relatively easy to estimate; \textbf{see the last paragraph of Section \ref{sec:accuracies} for details}.

\item \emph{Continuing on the numerical experiments, models 1 to 5 impose various structures on $\bs{\Sigma}$. It will be interesting to see the performance of the proposed method if a spiked covariance structure is imposed on $\bs{\Sigma}$.}

  We have added a spiked covariance matrix as model 6 in our new simulations. As mentioned in our response above, we found that models 5 and 6 constitute settings where our methods may not work as well as existing approaches.

\item \emph{Typo on page 6: "The density of $f(\cdot \mid \bs{\eta}_{jk})$ of ...".}

  This has been fixed.
\end{enumerate}

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-12-01 msg.tex
revised for biometrics to respond to reviewer comments

% \documentclass[12pt]{article}
\documentclass[useAMS,referee,usenatbib]{biom}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath, authblk}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
%\usepackage{graphics}
%\usepackage{afterpage}
%\usepackage{xcolor}
%\usepackage{citeppp}
%\usepackage[usenatbib]{natbib}
%\usepackage[round]{natbib}
%\usepackage{url} 
\usepackage{graphicx}
% \usepackage{subcaption}
\usepackage{xcolor}

@@

\label{firstpage}
\begin{abstract}
  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features. Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks. To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues. In this paper, we propose a new approach to estimating high-dimensional covariance matrices that can have better performance than existing methods. We first frame covariance matrix estimation as a compound decision problem. This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class. Simulation results and gene network inference in a single cell RNA-seq dataset show that our approach is comparable to or can outperform a number of state-of-the-art proposals in a wide range of settings.
\end{abstract}

@@

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications. However, in modern problems the number of features can be of the same order as or exceed the sample size. This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}. Accurate gene network estimation has become even more important with the advent of single-cell RNA-sequencing (scRNA-seq) technology, where interest now centers on inferring cell- or cell type-specific gene networks \citep{aibar2017scenic, dai2019cell, stegle2015computational}. For example, in Section \ref{gene analysis} we study cell type-specific gene network inference using scRNA-seq data from honey bee brains. \blue{Need to do analysis}

@@

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have lower estimation error than the sample covariance matrix. A first example was the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the sample covariance matrix toward a scaled identity matrix. More recently, nonlinear shrinkage methods were developed \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}. These shrink the eigenvalues of the sample covariance matrix in a data-adaptive fashion. Linear shrinkage can be viewed as a special case of nonlinear shrinkage, as it shrinks sample eigenvalues toward their global means.

Nonlinear shrinkage estimators have desirable optimality properties \citep{ledoit2018analytical} and show excellent performance. However, most existing nonlinear shrinkage estimators belong to the class of rotation-equivariant estimators \citep{bun2016rotational, stein1975estimation, stein1986lectures}, which modify the sample eigenvalue estimates but still use the sample eigenvectors. On the other hand, sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that certain classes of non-rotationally invariant estimators may outperform existing unstructured covariance matrix estimators.

@@

Here we propose a new class of unstructured non-rotationally invariant estimators for high-dimensional covariance matrices. Our approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}; see Section \ref{sec:compound}. This framing motivates us to vectorize the covariance matrix and solve the resulting vector estimation problem using a nonparametric empirical Bayes procedure, which has been shown in the compound decision literature to have excellent properties \citep{jiang2009general, koenker2014convex, efron2019bayes}. We then reassemble the estimated vector into matrix form and project onto the space of positive-definite matrices to give our final estimator. Surprisingly, though our vectorized approach essentially ignores the matrix structure, it can still substantially outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our compound decision approach is conceptually similar the empirical Bayes approach taken by \citet{dey2018corshrink} to estimate a high-dimensional correlation matrix. We discuss in detail the connections with their work in Section \ref{sec:proposed}. \blue{Need to add to numerical results}

The article is organized as follows. In Section \eqref{sec:method}, we briefly review compound decision theory and then introduce our proposed approach. In Section \eqref{numerical results} we illustrate the performance of our method in simulations and a gene expression dataset. Finally, Section \eqref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub. \blue{Is this implemented in the package yet?}

\section{\label{sec:compound}Compound decision problem formulation}
\subsection{\label{sec:background}Background}

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. A common measure of estimation performance is the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate. This paper constructs a new non-rotationally invariant $\bs{\delta}$ that in simulations and a data analysis has comparable or lower risk \eqref{frobenius risk} compared to existing methods.

Our proposed approach is motivated by the observation that minimizing \eqref{frobenius risk} is a type of compound decision problem, which we briefly review in this section. Introduced by \citet{robbins1951asymptotically}, compound decision problems study the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk
\begin{equation}
\label{compound risk}
R(\bs{\theta},\bs{\delta}) = \frac{1}{n}\sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))
\end{equation}
where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the homoscedastic Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.

A key property of compound decision problems is that while a given $Y_i$ or $\theta_i$ seems like it should offer no help for estimating $\theta_j$ when $j \ne i$, in fact borrowing information across all indices $i = 1, \ldots, n$ to estimate $\theta_j$ is superior to estimating each $\theta_j$ using the corresponding $Y_j$ alone. A classical example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}.

The discussion above shows that covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can be viewed as a compound decision problem. Furthermore, one special consequence of choosing \eqref{frobenius risk} as the risk measure is that the matrix estimation problem becomes equivalent to a vector estimation problem, specifically the problem of estimating every component of the vector $(\sigma_{11},\ldots,\sigma_{pp})^\top$ under average squared error loss. This interpretation is conceptually very similar to the classical Gaussian sequence problem.

We therefore propose to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation. One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive semidefinite. This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}. To resolve this issue, we project the result estimate into the space of positive-definite matrices; see Section \ref{posdef}.

\subsection{\label{sec:connections}Connections to existing work}

>>>>

% \documentclass[12pt]{article}
\documentclass[useAMS,referee,usenatbib]{biom}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath, authblk}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{comment}
\usepackage{lscape} 
%\usepackage{graphics}
%\usepackage{afterpage}
%\usepackage{xcolor}
%\usepackage{citeppp}
%\usepackage[usenatbib]{natbib}
%\usepackage[round]{natbib}
%\usepackage{url} 
\usepackage{graphicx}
% \usepackage{subcaption}
\usepackage{xcolor}
\usepackage{url}

@@

\label{firstpage}
\begin{abstract}
  Covariance matrix estimation is a fundamental statistical task in many applications, but the sample covariance matrix is sub-optimal when the sample size is comparable to or less than the number of features. Such high-dimensional settings are common in modern genomics, where covariance matrix estimation is frequently employed as a method for inferring gene networks. To achieve estimation accuracy in these settings, existing methods typically either assume that the population covariance matrix has some particular structure, for example sparsity, or apply shrinkage to better estimate the population eigenvalues. In this paper, we study a new approach to estimating high-dimensional covariance matrices. We first frame covariance matrix estimation as a compound decision problem. This motivates defining a class of decision rules and using a nonparametric empirical Bayes $g$-modeling approach to estimate the optimal rule in the class. Simulation results and gene network inference in an RNA-seq experiment in mouse show that our approach is comparable to or can outperform a number of state-of-the-art proposals, particularly when the sample eigenvectors are poor estimates of the population eigenvectors.
\end{abstract}

@@

Covariance matrix estimation is a fundamental statistical problem that plays an essential role in various applications. However, in modern problems the number of features can be of the same order as or exceed the sample size. This high-dimensional setting is especially common in genomics, where covariance matrices are used to model gene networks but the number of genes can be much larger than the number of biological replicates \citep{schafer2005shrinkage, markowetz2007inferring}. For example, in Section \ref{gene analysis} we study brain region-specific gene networks in mouse using bulk RNA-sequencing data.

@@

In contrast, unstructured methods do not make any assumptions about the population covariance matrix, yet still have lower estimation error than the sample covariance matrix. Many of these methods can be interpreted as rotationally invariant estimators, which shrink the sample eigenvalues in a data-adaptive fashion \citep{bun2016rotational, stein1975estimation, stein1986lectures}. One example is the linear shrinkage approach of \citet{ledoit2004well}, which shrinks the eigenvalues toward their global mean. Some more recently developed nonlinear shrinkage methods have been shown to be asymptotically optimal in this class \citep{ledoit2012nonlinear, ledoit2019quadratic, lam2016nonparametric}.

Though these optimal estimators perform extremely well, \citet{bun2016rotational} showed that the eigenvectors of any rotationally invariant estimate must be the same as those of the sample covariance matrix. This can be problematic because sample eigenvectors are not consistent when the dimension and the sample size increase at the same rate \citep{mestre2008asymptotic}. This suggests that unstructured but non-rotationally invariant covariance matrix estimators may be worth exploring, as these can modify both the sample eigenvectors as well as the eigenvalues. The difficulty is to identify a reasonable class of estimators to pursue.

@@

In this paper we define and explore one such class of estimators. This approach is based on interpreting the covariance matrix estimation problem as a compound decision problem \citep{robbins1951asymptotically}. Section \ref{sec:compound} shows how this framing motivates vectorizing the covariance matrix and solving the resulting vector estimation problem using nonparametric empirical Bayes procedures studied in the compound decision literature \citep{jiang2009general, koenker2014convex, efron2019bayes}. The vector estimate can then be reassembled into matrix form and projected into the space of positive-definite matrices to give a final covariance matrix estimator. Surprisingly, though this vectorized approach essentially ignores the matrix structure, it can sometimes outperform a number of state-of-the-art proposals in simulations and a real data analysis. Our results suggest that vectorized approaches may be preferrable when the sample eigenvectors are poor estimates of the population eigenvectors. Our compound decision approach is closely related to the correlation matrix estimator of \citet{dey2018corshrink}. We discuss in detail the connections with their work in Section \ref{sec:proposed}.

The article is organized as follows. In Section \ref{sec:method}, we briefly review compound decision theory and then introduce our proposed estimator. In Section \ref{numerical results} we illustrate its performance in simulations and a gene coexpression network inference problem. Finally, Section \ref{sec:discussion} concludes with a discussion. Our procedure is implemented in the R package \verb|cole|, available on GitHub at \url{http://github.com/sdzhao/cole}.

\section{\label{sec:compound}Covariance matrix estimation as a compound decision problem}
\subsection{\label{sec:background}Problem formulation}

Given $n$ observations $\bs{X}_1,\ldots,\bs{X}_n$ independently generated from a $p$-dimensional $\mathcal{N}(\bs{0},\bs{\Sigma})$, our goal is to find an estimator $\bs{\delta}(\bs{X})$ of $\bs{\Sigma}$. This section argues that covariance matrix estimation can be viewed as a compound decision problem when estimation performance is measured using the scaled squared Frobenius risk
\begin{equation}
\label{frobenius risk}
R(\bs{\Sigma}, \bs{\delta}) = \frac{1}{p^2} \sum_{j,k=1}^p \mathbb{E}[\{\delta_{jk}(\bs{X})-\sigma_{jk}\}^2],
\end{equation}
where $\sigma_{jk}$ is the $jk$th entry of $\bs{\Sigma}$ and $\delta_{jk}(\bs{X})$ is its corresponding estimate. This compound decision formulation will motivate a new class of unstructured non-rotationally invariant estimators.

Compound decision problems involve the simultaneous estimation of multiple parameters $\bs{\theta} = (\theta_1, \ldots, \theta_n)^\top$ given data $\bs{Y} = (Y_1, \ldots, Y_n)^\top$, with $Y_i\sim P_{\theta_i}$ \citep{robbins1951asymptotically}. Specifically, the goal is to develop a decision rule $\bs{\delta}(\bs{Y}) = (\delta_1(\bs{Y}),\ldots,\delta_n(\bs{Y}))$ that minimizes the compound risk $ n^{-1} \sum_{i=1}^n \bb{E} L(\theta_i,\delta_i(\bs{Y}))$, where $L$ is a loss function measuring the accuracy of $\delta_i(\bs{Y})$ as an estimate of $\theta_i$. A classical example is the classical Gaussian sequence problem, where $Y_i \sim N(\theta_i, 1)$ independently and $L(t, d) = (t - d)^2$ is the squared error loss \citep{johnstone2017gaussian}.

Covariance matrix estimation under the Frobenius risk \eqref{frobenius risk} can therefore be viewed as a compound decision problem where the goal is to simultaneously estimate the components of the vector
\begin{equation}
  \label{eq:sigma}
  \bs{\sigma} = (\sigma_{11}, \ldots, \sigma_{1p}, \ldots, \sigma_{p1}, \ldots,\sigma_{pp})^\top
\end{equation}
under average squared error loss.  One important difference between estimating a covariance matrix versus a vector is that the former should have additional structure, and in particular should be at least positive semidefinite. Notably, however, this structure is not incentived by the Frobenius risk \eqref{frobenius risk}. As a result, there can exist estimators of $\bs{\Sigma}$ that achieve low values of \eqref{frobenius risk} but which are not positive-definite. This will in fact be true of one of the estimators we propose in Section \ref{sec:proposed}. To resolve this issue, we can project any estimate into the space of positive-definite matrices; see Section \ref{posdef}.

The value of this formulation is that it allows us to apply ideas from the compound decision literature to covariance matrix estimation.  A key property of compound decision problems is that while information about index $i$ seems like it should offer no help for estimating a parameter with a different index $j$, in fact borrowing information across all indices is superior to considering each index alone. An example of this phenomenon is the James-Stein estimator \citep{james1961estimation}, which estimates $\bs{\theta} = E \bs{Y}$ in the Gaussian sequence problem by shrinking each $Y_i$ toward $0$ by a factor that depends on all components of $\bs{Y}$. \citet{james1961estimation} showed that when $n \geq 3$, the James-Stein estimator dominates the maximum likelihood estimator, which simply estimates $\bs{\theta}$ using $\bs{Y}$. A long line of subsequent work has led to much more sophisticated and accurate procedures for estimating $\bs{\theta}$ \citep{brown2009nonparametric, jiang2009general, johnstone2017gaussian, lindley1962discussion, fourdrinier2018shrinkage}. A natural next step is therefore to apply modern ideas for Gaussian sequence estimation to covariance matrix estimation, which we do in Section \ref{sec:method}.

\subsection{\label{sec:connections}Reinterpretation of existing methods}
